### 2017-05-15 assignments/2/paper.tex

\documentclass[12pt]{article}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath,bm}            % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{listings}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[]{algorithm2e}
\usepackage{parskip} 			   % no paragraph indentation

\usetikzlibrary{arrows,automata}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\lstset{ % General setup for the package
    language={[LaTeX]TeX},
    basicstyle=\footnotesize\sffamily,
    tabsize=4,
    columns=fixed,
    keepspaces,
    commentstyle=\color{red},
    keywordstyle=\color{blue},
    xleftmargin=.1\textwidth,
    xrightmargin=.1\textwidth
}

\begin{document}

\nocite{*}

\title{Randomized Algorithms \\
       Assignment 2}

\author{Viktor Hansen \& Simon Rueskov Schleicher}

\maketitle

\begin{abstract}
  This is the second weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
\end{abstract}

\pagebreak

\section*{Problems}

\subsection*{2.1}
We prove by induction over k that for any deterministic evaluation algorithm, there are instances of $T_{d,k}$ that forces the algorithm to read the values on all $d^{2k}$ leaves and the last leaf read determines whether the root is 1 or 0.\\
\\
Base step: for k=0 we only have one leaf, so any deterministic evaluation algorithm has to read them all. Furthermore that leaf is the root, so it determines whether the root is 1 or 0.\\
\\
Induction step: We assume it holds for any tree with size $k$, and need to show that it holds for any tree of size $k+1$. For a tree of size $k+1$, we have at the top a root that is an AND-note. Below that we have two OR-nodes, and below that we have 4 trees of size k. Since a deterministic evaluation algorithm looks at the nodes in a deterministic order, it also looks at the nodes in each of the subtrees in a specific order, and we can use the induction hypothesis.\\
There has to be one of the subtrees that are the first to get its last node inspected. We use the induction hypothesis and set the last node inspected in that subtree in the way that makes the root of that subtree 0. Since the parent to that root is an OR-node, we don't know yet whether it will be 1 or 0.\\
The 2nd subtree to get its last node inspected has 2 possibilities: Its root either shares a parent with the subtree we have already evaluated or it doesn't.\\
In the case where it shares a parent, we set the last node inspected in the way that makes the root 1. The parent OR-node can now look at a 1 and a 0, so the result is 1, but then we don't know the result of the AND-node in the root of the whole tree. We set the last node in the 3rd subtree to get fully evaluated in such a way that the root of that subtree is 0, and we have to inspect the 4th subtree. We can set the last node in the 4th subtree both so the root of the whole tree is 0 and so the root of the whole tree is 1.
In the case, where the 2nd subtree fully evaluated does not share a parent with the first, we set the last node in that subtree, so the root of the tree is 0, and we can't evaluate the parent OR-node. The 3rd subtree fully evaluated now either shares a parent with the first or the second subtree fully evaluated. We set the last node in the 3rd subtree in such a way that the root is 1, and we then have to also fully evaluate the 4th subtree before we can evaluate the root of the whole tree. Depending what we set the last node inspected in the 4th subtree, the root of the whole tree can both result in 0 or 1.\\
And we're done.

\subsection*{3.1}
Let $X$ be the number of empty bins after $m$ balls and $X_i$ be 1 if bin $i$ is empty after $m$ balls and 0 otherwise for $0\leq i\leq n$. For $X_i$ to be empty after m balls, we need all the m balls to land in a different bin, so we have $Pr[X_i]=\left(1-\frac{1}{n}\right)^m$ and $$E[X]=\Sigma_{i=0}^nE[X_i]=\Sigma_{i=0}^nPr[X_i]=\Sigma_{i=0}^n\left(1-\frac{1}{n}\right)^m=n\left(1-\frac{1}{n}\right)^m$$
For $m=n$ we get $$E[X]=n\left(1-\frac{1}{n}\right)^n\approx\frac{n}{e}$$

\subsection*{3.3(a)}
Let $X$ be the number of memory modules that has received exactly 1 memory request after $n$ memory requests has been sent and $X_i$ be 1 if memory module $i$ has received exactly 1 memory request after $n$ memory requests has been sent and 0 otherwise for $0\leq i\leq n$. We have $Pr[X_i]={n\choose 1}\left(\frac{1}{n}\right)^1\left(1-\frac{1}{n}\right)^{n-1}=\left(1-\frac{1}{n}\right)^{n-1}$.\\
Let $Y$ be the number of memory modules that has received exactly 2 memory request after $n$ memory requests has been sent and $Y_i$ be 1 if memory module $i$ has received exactly 2 memory request after $n$ memory requests has been sent and 0 otherwise for $0\leq i\leq n$. We have $Pr[Y_i]={n\choose 2}\left(\frac{1}{n}\right)^2\left(1-\frac{1}{n}\right)^{n-2}={n\choose 2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}$.
Let $Z$ be the number of processors whose requests are satisfied. We have \begin{align*}E[Z]&=E[X]+2E[Y]=\left(\Sigma_{i=0}^nE[X_i]\right)+2\Sigma_{i=0}^nE[Y_i]=\left(\Sigma_{i=0}^nPr[X_i]\right)+2\Sigma_{i=0}^nPr[Y_i]\\&=\left(\Sigma_{i=0}^n\left(1-\frac{1}{n}\right)^{n-1}\right)+2\Sigma_{i=0}^n{n\choose 2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}\\&=n\left(1-\frac{1}{n}\right)^{n-1}+2n\frac{n(n-1)}{2}\frac{1}{n^2}\left(1-\frac{1}{n}\right)^{n-2}=\frac{n}{1-\frac{1}{n}}\left(1-\frac{1}{n}\right)^n+(n-1)\left(1-\frac{1}{n}\right)^{n-2}\\&=\frac{n-1}{\left(1-\frac{1}{n}\right)^2}\left(1-\frac{1}{n}\right)^n+\frac{n-1}{\left(1-\frac{1}{n}\right)^2}\left(1-\frac{1}{n}\right)^n\approx\frac{2e(n-1)}{\left(\frac{n-1}{n}\right)^2}=\frac{2en^2}{n-1}\end{align*}
This is around $2en$, which is of course impossible (we can satisfy more processor than are available), so we must have made a mistake somewhere.

\pagebreak

\section*{Summary}
\subsection*{Balls into bins}
Given $n$ balls and $n$ bins, and ball $i$ goes into bin $j$ with probability $1/n$. Then the expected number of balls in bin $j$ is 1. Let $\varepsilon^=_j(k)$ denote the event where bin $j$ receives exactly $k$ balls. Then
\begin{align*}
\mathbf{Pr}\left[ \varepsilon^=_j(k) \right] = \binom{n}{k}\frac{1}{n}^n \left( 1-\frac{1}{n} \right)^{n-k} \leq \binom{n}{k}\frac{1}{n}^n \leq \binom{ne}{k}^k \left( \frac{1}{n} \right)^n = \left(\frac{e}{k}\right)^k 
\end{align*}
Now, let $\varepsilon^\geq_j(k)$ denote the event where bin $j$ receives $k$ or more balls, then
\begin{align*}
\mathbf{Pr}\left[ \varepsilon^\geq_j(k) \right] = \sum_{i=k}^{n} \mathbf{Pr}\left[ \varepsilon^=_j(i) \right] \leq \sum_{i=k}^{n} \left(\frac{e}{i}\right)^i \leq \sum_{i=k}^{n} \left(\frac{e}{k}\right)^i \leq \left(\frac{e}{k}\right)^k \sum_{i=0}^{\infty} \left( \frac{e}{k} \right)^i = \left( \frac{e}{k} \right)^k \frac{1}{1-\frac{e}{k}}
\end{align*}
Choosing $k^*=\frac{3 \ln n}{\ln \ln n}$ and inserting it into the above yields
\begin{align*}
\mathbf{Pr}\left[ \varepsilon_j^\geq(k^*) \right] \leq \frac{1}{n^2}.
\end{align*}
The probability of any bin receiving more than $k^*$ balls is given by
\begin{align*}
\mathbf{Pr}\left[ \cup_{j=1}^{n} \varepsilon_j^\geq(k^*) \right] \leq \sum_{j=1}^n \mathbf{Pr}\left[ \varepsilon_j^\geq(k^*) \right] \leq n \frac{1}{n^2} = \frac{1}{n}
\end{align*}

\subsection*{Markov's Inequality}
Given a non-negative random variable $Y$ and $t > 0$, then
\begin{align*}
\mathbf{Pr}\left[ Y \geq t \mathbf{E}\left[ Y \right] \right] \leq \frac{1}{t} \;\; \mathrm{ and } \;\; \mathbf{Pr}\left[ Y \geq t \right] \leq \frac{\mathbf{E}\left[ Y \right]}{t}
\end{align*}
To show the rightmost form, we get
\begin{align*}
\mathbf{E}\left[ Y \right] = \sum_{i} i \cdot \mathbf{Pr}\left[ Y=i \right] \geq \sum_{i \geq t} i \cdot \mathbf{Pr}\left[ Y=i \right] \geq t \cdot \sum_{i \geq t} \mathbf{Pr}\left[ Y=i \right] \geq t \cdot \mathbf{Pr}\left[ Y \geq i \right] 
\end{align*}
The leftmost form follows by substituting $t$ by $t \cdot \mathbf{E}\left[ Y \right]$ into the rightmost form.

\subsection*{Chebyshev's Inequality}
Given rand. var. $X$, define the variance of $X$, denoted by $\sigma^2$, as $\sigma^2 = \mathbf{E}\left[ \left( X - \mu_X \right)^2 \right]$, where $\mu_X = \mathbf{E} \left[ X \right]$. Now, given a random variable $X$ and $t > 0$, then
\begin{align*}
\mathbf{Pr}\left[ \left| X-\mu_X \right| \geq t\sigma_X \right] &= \mathbf{Pr}\left[ \left( X-\mu_X \right)^2 \geq t^2\sigma_X^2 \right] \\
&= \mathbf{Pr}\left[ \left( X-\mu_X \right)^2 \geq t^2 \mathbf{E}\left[\left( X-\mu_X \right)^2\right] \right] \\
&= \mathbf{Pr}\left[ Y \geq t^2\mathbf{E}\left[Y\right] \right] \\
&= \frac{1}{t^2}
\end{align*}
The last step follows from applying Markov's inequality.

\subsection*{Lazy Selection}
Refer to the algorithm given in the book. Let $S(k)$ denote the $k$th smallest element in $S$ and $r_{S(a)}$ denote the rank of $a$ in $S$. Any known det. alg. requires $\geq 3n$ comparisons to identify $S_{k}$. We want to show that \textsc{LazySelect} finds $S(k)$ with $2n + O(n)$ comparisons with prob. $1-O(n^{-1/4})$. The algorithm can err. if $S(k) < a$. For this to happen, fewer than $\mathcal{l}$ samples are smaller than $S(k)$ for $i=1, \hdots, n^{3/4}$, $X_i=1$ if $i$th sample $\leq S(k)$, and $X_i = 0$ otherwise. Let $X=\sum^{n^{3/4}}_{i=1} X_i$. We will bound the error probability using Chebyshev's inequality. We get $\mu_X = \sum^{n^{3/4}}_{i=1} \mu_{X_i} = n^{3/4}\cdot \frac{k}{n}=kn^{-1/4}$ and $\sigma_X^2 = \sum^{n^{3/4}}_{i=1} \sigma^2_{X_i} = n^{3/4}\frac{k}{n}\left( 1-\frac{k}{n} \right) \leq \frac{1}{4}n^{3/4}$ implying $\sigma_X \leq \frac{1}{2}n^{3/8}$ and
\begin{align*}
\mathbf{Pr} \left[ X \leq l \right] &= \mathbf{Pr} \left[ X \leq \mu_X - \sqrt{n} \right] = \mathbf{Pr} \left[ \mu_X - X \geq \sqrt{n} \right] \leq \mathbf{Pr} \left[ \left| \mu_X - X \right| \geq \sqrt{n} \right] \\
&\leq \mathbf{Pr} \left[ \left| \mu_X - X \right| \geq 2n^{1/8}\sigma_X \right] \leq \frac{1}{4n^{1/4}} = O(n^{-1/4})
\end{align*}
and hence the success probability is $\geq 1-O(n^{-1/4})$.

\subsection*{Two-point Sampling}
\begin{lem}
Let $a,b \in \mathbb{Z}_n$ where $n$ is a prime be picked independently and uniformly at random. Let $y_i = (a_i + b) \mod n$ for $i = 0, \hdots, n-1$. Then for all distinct $i,j \in \mathbb{Z}_n$, $y_i$ and $y_j$ are pairwise ind. and uniformly distributed in $\mathbb{Z}_n$. \\
\end{lem}

\begin{dfn}
Variables $X_1, \hdots, X_n$ are pairwise ind. iff. for all distinct $i,j$ and all $x_y$, $\mathbf{Pr}\left[ X_i = x \; | \; X_j = y \right] = \mathbf{Pr}\left[ X_i = x \right]$. \\
\end{dfn}

\begin{lem}
If $X_1, ... X_n$ are pairwise independent and $X = \sum X_{i=1}^n X_i$ then $\sigma^2_X = \sum_{i=1}^n \sigma_{X_i}^2$
\end{lem}

Given an algorithm $A(x,r)$ for deciding $L$ where $x \in L$ and $r$ is a random number in $\mathbb{Z}_n = \left\{ 0, \hdots, n-1 \right\}$ where $n$ is a prime s.t.
\begin{align*}
&\text{1) If } x \not\in L \text{ then } A(x,r) = 0 \; \forall r \in \mathbb{Z}_n \\
&\text{2) If } x \in L \text{ then } A(x,r) = 1 \text{ for at least half of the choices of } r \in \mathbb{Z}_n
\end{align*}
Simple method: Pick $t$ numbers, $r_i, \hdots, r_t$ independently and uniformly at random from $\mathbb{Z}_n$. Output $\lor_{i=1}^t A(x,r_i)$. Given $x \in L$ the failure probability $\leq 1/2^t$. For $t=2$, the failure probability $\leq 1/4$.

Two-point method: Now let $a,b \in \mathbb{Z}_n$ be picked uniformly at random. Let $r_i = (a_i + b) \mod n$, $Y_i = A(x, r_i)$ for $i = 0, \hdots, n-1$ and $Y=\sum_{i=0}^{n-1}Y_i$. We wish to bound $\mathbf{Pr}\left[ Y=0 \right]$ using Chebyshev's inequality. Now $\mu_{Y} = \sum^{n-1}_{i=0}\mu_{Y_i} \geq n/2$ and $\sigma_Y^2 = \sum_{i=0}^{n-1} \sigma^2_{Y_{i}} \geq n/4 \Rightarrow \sigma_Y \geq \sqrt{n}/2$. Inserting into Chebyshev's Inequality yields
\begin{align*}
\mathbf{Pr}\left[ Y = 0 \right] \leq \mathbf{Pr}\left[ \left| Y - \mu_Y \right| \geq \sqrt{n} \sigma_Y \right] \leq \frac{1}{n}
\end{align*}
which is better than $1/4$ for $n \geq 4$.

\subsection*{Coupon Collector's Problem}
Given $n$ types of coupons, a trial means picking a coupon independently and at random. Let $X = \text{\# of trials to get all } n \text{ coupon types}$. For $i=0, \hdots, n-1$, the $i$th epoch consists of the trials starting just after the $i$th success and ending in the $(i+1)$th success. Define $X_i = \text{\# of trials in the } i\text{th epoch.}$, then $X=\sum_{i=0}^{n-1} X_i$. Now
\begin{align*}
\mathbf{E}\left[ X \right] = \sum_{i=0}^{n-1} \mathbf{E}\left[ X_i \right] = \sum_{i=0}^{n-1} \frac{n}{n-i} = \sum_{i=1}^{n} \frac{n}{i} =  n \sum_{i=1}^{n} \frac{1}{i} = n H_{n} = n \ln n + O(n)
\end{align*}
Let $\varepsilon_i^r$ denote the event that coupon type $i$ is not picked in the first $r$ trials. Then
\begin{align*}
\mathbf{Pr}\left[ \varepsilon_i^r \right] = \left(1-\frac{1}{n}\right)^r \leq e^{-r/n}
\end{align*}
Pick $r=\beta n \ln n$, where $\beta$ is a constant, then $\mathbf{Pr}\left[ \varepsilon_i^r \right] \leq e^{-\beta \ln n} = n^{-\beta}$ and
\begin{align*}
\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq \sum_{i=1}^n \mathbf{Pr}\left[ \varepsilon_i^r \right] \leq n^{1-\beta}
\end{align*}
For $\beta=2$, we get $\mathbf{Pr}\left[ \cup_{i=1}^n \right] \leq 1/n$.

\subsection*{Chernoff Bounds}
Given $n$ independent indicator variables $X_1, \hdots, X_n$, $p_i = \mathbf{Pr}\left[ X_i = 1 \right]$, $0 < p_i < 1$ for $i=1, \hdots, n$. Let $X=\sum_{i=1}^n X_i$ and $\mu = \mathbf{E} \left[ X \right] = \sum_{i=1}^n p_i$. Then
\begin{align*}
\mathbf{Pr}\left[ X > (1-\delta)\mu_X \right]
&= \mathbf{Pr}\left[ e^{tX} > e^{t(1-\delta)\mu_X} \right]
< \frac{\mathbf{E}\left[ e^{tX} \right]}{e^{t(1-\delta)\mu_X}}
= \frac{\prod_{i=1}^n \mathbf{E}\left[ e^{tX_i} \right]}{e^{t(1-\delta)\mu_X}} \\
&= \frac{\prod_{i=1}^n \left( p_i e^t + 1-p_i \right)}{e^{t(1-\delta)\mu_X}}
= \frac{\prod_{i=1}^n \left( 1 + p_i (e^t - 1) \right)}{e^{t(1-\delta)\mu_X}} \\
&\leq \frac{\prod_{i=1}^n e^{p_i (e^t-1)}}{e^{t(1-\delta)\mu_X}}
\leq \left( \frac{e^{(e^t-1)}}{e^{t(1-\delta)}} \right)^{\mu_X}
= \left( \frac{e^{\delta}}{\left( 1 + \delta \right)^{1+\delta}} \right)^{\mu_X}
\end{align*}
Furthermore
\begin{align*}
\mathbf{Pr}\left[ X < (1-\delta)\mu_X \right]
&= \mathbf{Pr}\left[ -X < -(1-\delta)\mu_X \right]
= \mathbf{Pr}\left[ e^{-tX} < e^{-t(1-\delta)\mu_X} \right] \\
&< \left( \frac{e^{-\delta}}{\left( 1-\delta \right)^{1-\delta}} \right)^\mu
= \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right) \ln (1-\delta)}} \right)^\mu \\
&< \left( \frac{e^{-\delta}}{e^{\left( 1-\delta \right)\left( -\delta - \frac{\delta^2}{2} \right)}} \right)^\mu
=  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2 + \frac{\delta^3}{2}}  } \right)^\mu \\
&<  \left( \frac{e^{-\delta}}{e^{-\delta - \frac{\delta^2}{2} + \delta^2}  } \right)^\mu = \left( \frac{1}{e^{-\frac{\delta^2}{2} + \delta^2}  } \right)^\mu
= \left( e^{\frac{\delta^2}{2} - \delta^2} \right)^\mu \\
&= e^{\mu\frac{\delta^2}{2} - \mu\delta^2} = e^{-\mu\sigma^2/2}
\end{align*}

\subsection*{Set-Balancing}
Given a matrix $A \in \left\{ 0,1 \right\}^{n \times n}$, find a vector $\overline{b} \in \left\{ -1, 1 \right\}^n$ s.t. $|| A\overline{b} ||_{\infty}$ is minimized. Pick each entry of $\overline{b}$ to be $1$ with probability $1/2$ and $-1$ otherwise. Now, $\mathbf{E} \left[ || A\overline{b} ||_{\infty} \right] = 0$. Consider a row $i$, $A_{i}$, of $A$, and assume for now that all entries are 1. Let $X$ be the \# of 1's in $\overline{b}$. Now
\begin{align*}
\mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] &\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \; \lor \; X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
&\leq \mathbf{Pr} \left[ X > \frac{n}{2} + \sqrt{2n \ln n} \right] + \mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right] \\
&\leq \frac{1}{n^2} + \frac{1}{n^2} = \frac{2}{n^2}
\end{align*}
and the third inequality follows from
\begin{align*}
\mathbf{Pr} \left[ X < \frac{n}{2} - \sqrt{2n \ln n} \right]
&= \mathbf{Pr} \left[ X < \left( 1 - \frac{2\sqrt{2n \ln n}}{n} \right)\frac{n}{2} \right]
< e^{-\frac{n}{2} \left( \frac{2\sqrt{2n \ln n}}{n}\right)^2\frac{1}{2}} \\
&< e^{-\frac{n}{4} \left( \frac{8n \ln n}{n^2}\right)} = e^{-2\ln n} = 1/n^2
\end{align*}
where the first inequality follows from applying the second Chernoff bound to the inequality with $\delta=\frac{2\sqrt{2n \ln n}}{n}$ and $\mu=n/2$. Finally applying an union bound over all rows we get
\begin{align*}
\mathbf{Pr} \left[ || A\overline{b} ||_{\infty} > 2\sqrt{2n \ln n} \right] \geq \sum_{i=1}^n \mathbf{Pr} \left[ | A_i\overline{b} | > 2\sqrt{2n \ln n} \right] \leq \sum_{i=1}^n \frac{2}{n^2} = \frac{2}{n}
\end{align*}

\end{document}