### 2017-09-07 assignments/5/paper.tex

\documentclass[12pt]{article}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath,bm}            % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{listings}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[]{algorithm2e}
\usepackage{parskip} 			   % no paragraph indentation

\usetikzlibrary{arrows,automata}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{alg}[thm]{Algorithm}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\lstset{ % General setup for the package
    language={[LaTeX]TeX},
    basicstyle=\footnotesize\sffamily,
    tabsize=4,
    columns=fixed,
    keepspaces,
    commentstyle=\color{red},
    keywordstyle=\color{blue},
    xleftmargin=.1\textwidth,
    xrightmargin=.1\textwidth
}

\begin{document}

\nocite{*}

\title{Randomized Algorithms \\
       Assignment 5}

\author{Viktor Hansen \& Simon Rueskov Schleicher}

\maketitle

\begin{abstract}
  This is the fifth weekly assignment for the Randomized Algorithms course offered at The Department of Computer Science, Uni. Copenhagen.
\end{abstract}

\pagebreak

\section*{5.3)}
\paragraph{a)} Consider a graph $G$ with $n$ vertices and $nd/2$. Delete each vertex and its incident edges with probability $1-1/d$. Let $N_i=1$ if vertex $i$ is deleted and $N=\sum_{i=1}^{n} N_i$ be the total number of vertices after all deletions. Then
\begin{align*}
\mathbf{E}[N] = \sum_{i=1}^{n} \mathbf{E}[N_i] = \sum_{i=1}^{n} 1/d = n/d
\end{align*}
In a similar fashion, let $E_i=1$ if edge is is not deleted. Then
\begin{align*}
\mathbf{E}[E] &= \sum_{i=1}^{nd/2} \mathbf{E}[E_i] = \sum_{i=1}^{nd/2} 1-\mathbf{Pr}\left[ \text{edge } i \text{ belonging to a deleted node.} \right] \\
&= \sum_{i=1}^{nd/2} 1-(1-1/d^2) \\
&= \frac{n}{2d}
\end{align*}

\paragraph{b)} For any graph $G=(V,E)$, a trivial lower bound on the size of any independent set is $\max\left\{0,|V| - |E|\right\}$, since each edge can reduce the size of such an independent set by at most 1. Let $X$ and $Y$ denote the number of vertices and edges in $G$ after the deletion in a). Now, repeatedly pick a vertex, $v$, of $G$ and remove all its neighbours in $G$ remain and add $v$  to an initially empty set, $S$, until no vertices remain in $G$. Now $S$ is clearly an independent set, and its size is bounded from below by $X-Y$. Now
\begin{align*}
\mathbf{E}\left[ S \right] &\geq \mathbf{E}\left[ \max\left\{0,X-Y\right\} \right] = \mathbf{E}\left[ X \right] - \mathbf{E}\left[ Y \right] = \frac{n}{d} - \frac{n}{2d} = \frac{n}{2d}
\end{align*}
Since $\mathbf{E}\left[ S \right] \geq n/2d$, $S$ achieves a value of at least $n/2d$ with positive probability, by (2) of the probabilistic method. By (1) of the probabilistic method, this implies that an independent set of size at least $n/2d$ must exist.

\section*{2) Show that lemma 5.11 implies corollary 5.12}
Let $x_i = \frac{1}{d+1}$ for all $1 \leq i \leq n$. Then
\begin{align*}
x_i\prod_{(i,j) \in E} (1-x_j) &= \frac{1}{d+1} \left( 1-\frac{1}{d+1} \right)^{\deg(i)} \\
&\geq \frac{1}{d+1} \left( 1-\frac{1}{d+1} \right)^{d} \\
&\geq \frac{1}{d+1} \cdot \frac{1}{e} \geq p \geq  \mathbf{Pr} \left[ \varepsilon_i \right]
\end{align*}
By Lovasz Local Lemma,
\begin{align*}
\mathbf{Pr} \left[ \cap_{i=1}^{n} \overline{\varepsilon}_i  \right] \geq \prod_{i=1}^n \left( 1 - \frac{1}{d+1} \right) \geq \left( 1 - \frac{1}{d+1} \right)^n > 0
\end{align*}
when $\frac{1}{(d+1)e} \geq p \Leftrightarrow 1 \geq ep(d+1)$.

\section*{3) 5.13}
Using an approach similar to that of section 5.6, we adopt a computation-tree view of the algorithm. Each level of the tree corresponds to assigning a vertex to $A$ or $B$. If for any step, the algorithm can assign a vertex to $A$ or $B$ s.t. $\mathbf{Pr}\left[\text{\# edges crossing} < m/2\right] < 1$, the algorithm must terminate with $\text{\# edges crossing} \geq m/2$ at a leaf node. Theorem 5.1 tells us that this is the case at the root, so if we can maintain this invariant for each step, it is guaranteed to find a solution deterministically. Let $C_i, A_i, B_i$ and $U_i$ denote the size of the cut, $A$, $B$ and the set of unassigned vertices at each step $i$, for $1 \leq i \leq n$. Then $\mathbf{E}\left[ C_i \right] = \frac{1}{2} \mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{i+1} \right] + \frac{1}{2} \mathbf{E} \left[ C_{i+1} | v_{i+1} \in B_{i+1} \right]$ by employing the randomized partitioning strategy. It follows that $\mathbf{E}\left[ C_i \right] \leq \max \left\{ \mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{b+1} \right], \mathbf{E} \left[ C_{i+1} | v_{i+1} \in B_{i+1} \right] \right\}$ and thus choosing the assignment maximising the expectation can only decrease the probability of failure. The conditional expectation $\mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{i+1} \right]$ can be computed as follows:
\begin{align*}
\mathbf{E} \left[ C_{i+1} | v_{i+1} \in A_{i+1} \right] &= \left| A_{i+1} \times B_{i+1} \right| + \frac{1}{2} \left| U_{i+1}^2 \right| + \frac{1}{2} \left| A_{i+1} \times U_{i+1} \right|  + \frac{1}{2} \left| U_{i+1} \times B_{i+1} \right|\\
&= \left| A_{i+1} \times B_{i+1} \right| + \frac{1}{2} |\overline{A_{i+1} \times B_{i+1}}|
\end{align*}
and similarly for $\mathbf{E} \left[ C_{i+1} | v_{i+1} \in B_{i+1} \right]$. Since $\mathbf{E}\left[ C_i \right] \leq \mathbf{E}\left[ C_{i+1} \right]$ for all $1 \leq i \leq n$, we conclude that the the derandomised algorithm must produce a cut of size at least $m/2$.
\pagebreak

\section*{Summary May 18th and 23rd}
\subsection*{The Probabilistic Method}
\begin{enumerate}
\item If an object drawn from a universe has some property P with probability $> 0$, then some object in the universe has property P.
\item Any random var. achieves a value of at least (resp. at most) its expected value with positive probability.
\end{enumerate}

\subsection*{Max-Cut}
Given an undirected graph $G=(V,E)$, find a cut $V_1 \cup V_2$ of $V$ s.t. $\left\{ (u,v) \in E \; | \; u \in V_2, v \in V_2 \right\}$ is maximized. \\

\begin{alg}
For each $v \in V$, add it to $V_1$ w. prob. $1/2$, and add it to $V_2$ otherwise. \\
\end{alg}

\begin{thm}
$G$ has a cut with at least $m/2$ edges, where $m=|E|$.
\paragraph{Proof:} $\mathbf{E}[\text{\# edges crossing }(V_1,V_2)\text{ with above algo.}]=m/2$. Then by TPM, there exists a cut of $G$ with $\geq m/2$ edges crossing it.
\end{thm}

\subsection*{Max-SAT}
Given boolean formula in CNF with $m$ clauses, e.g. $(x_1 \lor \neg x_2 \lor x_3) \land (x_4 \lor x_1) \land \hdots$, and assume that no clause contains more than one occurrence of each variable. Then \\
\begin{thm}
Any B. formula in CNF with $m$ clauses each with at least $k$ literals has an assignment that satisfies at least $m(1-2^k)$ clauses. \\
\end{thm}

\begin{alg}
Set each variable to true with prob. $1/2$. \\
\end{alg}
Now, $\mathbf{Pr}[C \text{ sat. by assignment from algo.}] \geq 1-\frac{1}{2^k}$ and thus $\mathbf{E}[\text{\# clauses sat.}] \geq m(1-2^{-k})$, and the theorem follows by THM.

\paragraph{Method 1} The above is a $1/2$ approximation, i.e. $\geq m(1-2^{-k})$ clauses in expectation where each clause has $\geq k$ literals.

\paragraph{Method 2} Is a $\beta_k$-approximation. Let $C_i^-$, $C_i^+$ be the sets of negated and remaining variables in clause $i$, respectively. ILP formulation:
\begin{align*}
\max & \sum_{\text{clause } i} z_i\\
\text{s.t. }& \sum_{j \in C^+_i} y_j + \sum_{j \in C_i^-}(1-y_j) \geq z_i \; \forall \text{ clauses } i \\
\text{where }& y_j,z_i \in \left\{ 0,1 \right\} \; \forall i,j
\end{align*}

\begin{alg}
Solve linear relaxation of ILP where $\hat{y}_j, \hat{z}_i$ are fractional solutions and $\overline{y}_j$ is the rounded solution, s.t. $\mathbf{Pr}[\overline{y}_j = 1] = \hat{y}_j$. \\
\end{alg}

Consider a clause $i$ with $k$ literals. Assume that $C_i^- = \emptyset$. Then
\begin{align*}
\mathbf{\text{clause } i \text{ sat.}} &=1-\prod_{j \in C_i^+}(1-\hat{y}_j) \\
&\geq 1-\prod_{j \in C_i^+}\left(1-\frac{\hat{z}_j}{k}\right)=1-\left(1-\frac{\hat{z}_j}{k}\right)^k \\
&\geq\beta_k\hat{z}_i\text{, where } \beta_k=1-\left(1-\frac{1}{k}\right)^k
\end{align*}
Then $\mathbf{Pr}[\text{\# clauses sat. assuming every clause has at least } k \text{ literals}] \geq \sum_{\text{clause } i}\beta_k \cdot OPTLP \geq \beta_k \cdot OPT$. Thus $\sum \hat{z}_i \geq OPT$. Now, let $n_i = \mathbf{E}[\text{\# clauses sat. by method } i]$. Then \\
\begin{thm}
$\max \left\{ n_1, n_2 \right\} \geq \frac{3}{4} \cdot OPT$
\paragraph{Proof:}
\begin{align*}
\max \left\{ n_1, n_2 \right\} &\geq \frac{n1+n2}{2} \geq \sum_k \sum_{i \in S_k} \frac{(1-2^{-k}) + \beta_k}{2} \hat{z}_i \\
&\geq \sum_k \sum_{i \in S_k} \frac{3}{4} \hat{z}_i = \frac{3}{4} \sum_{i \in S_k}  \hat{z}_i \geq \frac{3}{4} \cdot OPT
\end{align*}
\end{thm}

\subsection*{OR-concentrator}
\begin{dfn}
An $(n,d,\alpha,c)$ OR-concentrator is a bipartite graph $G=(L,R,E)$ with $|L|=|R|=n$ s.t.
\begin{enumerate}
\item Every vertex in $L$ has deg. $\leq d$ and 
\item $\forall S \subseteq L : |S| \leq \alpha n \Rightarrow |N(S)| \geq c |S|$ \\
\end{enumerate}
\end{dfn}

\begin{thm}
Given $n \in \mathbb{N}, \; \exists (n,18,\frac{1}{3},2)$ OR-conc.
\paragraph{Proof:} For each $v \in L$, rand. algo picks $d$ neighbours in $R$, uniformly at random and with replacement. Connect $v$ with edges to its chosen neighbours.

Let $\varepsilon_S : \exists S \subseteq L, |S|=s, |N(S)| \leq cs$. Then
\begin{align*}
\mathbf{Pr}[\varepsilon_S]&=\binom{n}{s}\binom{n}{cs}\binom{cs}{n}^{ds} \\
&\leq \binom{ne}{s}^s\binom{ne}{cs}^cs\binom{cs}{n}^{ds} \\
&= \left( \binom{ne}{s}\binom{ne}{2s}^2\binom{2s}{n}^{18} \right)^s  \\
&= \left( n^{-15} s^{15} e^3 2^{16} \right)^5 \\
&\leq \left( n^{-15} \left( \frac{1}{3^n} \right)^{15} e^3 2^{16} \right)^5 \\
&= \left( \left( \frac{1}{3} \right)^{-15} e^3 2^{16} \right)^5 < \left( \frac{1}{2} \right)^s
\end{align*}
And $\mathbf{Pr}\left[ \cup_{S=1}^{\alpha n} \varepsilon_S \right] \leq \sum_{S=1}^{\alpha n} \left(\frac{1}{2}\right)^S < 1$. By TPM, an OR-concentrator exists with positive probability.
\end{thm}

\subsection*{Probability amplification}
Given an algorithm $A(x,t)$ where if $x \not\in L$ then $A(x,r) = 0$ $\forall r$ and if $x \in L$ then $A(x,r)=1$ for at least half the choices of $r$. \\

\begin{thm}
For suff. large $n \in \mathbb{N}$, there exists a bipartite graph $G=(L,R,E)$ s.t. $|L| = n$, $|R| = 2^{\log^2 n}$ and
\begin{enumerate}
\item every vertex in $R$ has deg. $\leq 12 \log^2 n$
\item $\forall S \subseteq L, |S|=\frac{n}{2} \Rightarrow |N(S)| \geq 2^{\log^2 n} - n$
\end{enumerate}
\paragraph{Proof: } For each $u \in L$, rand. algo picks $d=\frac{2^{\log^2 n}(4 \log^2 n)}{n}$ neighbours in $R$ uniformly and independently at random w. replacement. This gives a bipartite graph $G=(L,R,E)$. Let $v \in R$. We wish to apply Chernoff bound, so $\mathbf{E}[\deg(v)]=nd\frac{1}{2^{log^2 n}} = 2 \log^2 n$ and
\begin{align*}
\mathbf{Pr}\left[ \deg(v) > 3 \cdot 3 \log^2 n \right] < \left( \frac{e}{1+\delta}\right)^{1+\delta}\mu=\left( \frac{e}{3} \right)^{12\log^2 n}
\end{align*}
Taking the union bound over the events yields $$\mathbf{Pr}\left[ \bigcup_{v \in R} \left\{ \deg(v) > 12 \log^2 n \right\} \right] < 2^{\log^2 n} \cdot \left(\frac{e}{3} \right)^{12 \log^2 n} \leq \frac{1}{2}$$
Showing that the probability that condition 1 fails is at most 1/2. For condition 2:
\begin{align*}
\mathbf{Pr}[\text{cond. 2 fails}] &\leq \binom{n}{n/2}\left( \frac{2^{\log^2 n}}{n} \right)\left( 1- \frac{n}{2^{\log^2 n}} \right)^{\frac{dn}{2}} \\
&\leq \left( \frac{ne}{n/2}^{n/2} \right) \left( \frac{2^{\log^2 e}}{n} \right)^n e^{-\frac{dn}{2}\frac{n}{2^{\log^2 n}}} \\
&= (2e)^{n/2}\left( \frac{2^{\log^2 n}e}{n} \right)^n e^{-2n \log^2 n} \\
&< \frac{1}{2}
\end{align*}
Finally, by union bound we get $\mathbf{Pr}[\text{cond. 1 fails or cond. 2 fails}] < \frac{1}{2} + \frac{1}{2} = 1$. Theorem follows from TPM.

\end{thm}
To amplify prob, we can, with $\log^2 n$ random bits, pick a vertex $v \in R$ uniformly at random. For each $u \in N(v)$, run $A(x,r_u)$ where $r_u$ is the index of $u$. Output $\lor_{u \in N(v)} A(x,r_u)$. This yields a failure prob. $\leq \frac{n}{2^{\log^2 n}}$

\end{document}